---
title: "WUP_04"
author: "Jaelynn Schultz"
date: "Summer 2022"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)

mydata<-read_csv(file = "GSScsv.csv")

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(mydata,aes(x=DEGREE,y=WORDSUM))
```
```{r}
basicCN + geom_boxplot()
```

Based on this plot, it seems that the graduate level degree shows the highest score in comparison to the others. This is shown by the fact that the median is just about as high as the Q3 for both the bachelor's and junior college degrees. The LT High School's median is the lowest showing a median that is lower than the regular high school level by a small margin.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=mydata)
```

The mean descends from bachelor, graduate, high school, junior college, and then LT high school. Population must be taken into account here as we can see that the Bachelor's level has a population of 44 while the junior college only has 12 at the lowest. The regular high school level sits at the highest population with 85. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
```

Given that the p-value is well below 0.05 (0.000236), we can conclude that there is a significant difference between the categories of data.

```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 15.7, 2, 2))
plot(TukeyHSD(model1),las=2)
```

Based on the confidence intervals, we can clearly see that the high school and graduate levels are vastly different. We can see a similar difference between the LT high school and graduate degrees. It would appear the greatest differences are between high school degrees and graduate degrees. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

After reviewing the graphical, numerical, and inferential results, it is clear that there is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE. The graduate level shows the greatest scores on the WORDSUM test and those in the high school and LT high school levels show the worst scores. The greater one's degree, it would appear the better they begin to do on the WORDSUM test. 
